RED WINE QUALITY by SAMMIT RANADE
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(gridExtra)
library(dplyr)
library(reshape2)
library(GGally)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
#setwd("~/Desktop")
#getwd()
wr <- read.csv('wineQualityReds.csv' )
#View(wr)
names(wr)
summary(wr)
```

From the above summary, we can see from the statistics how each object varies
from Minimum to Maximum and their means, medians,etc.

The dataset I am planning to explore is the Red Wine Quality 
dataset. The guiding question for this dataset is "Which chemical properties 
influence the quality of red wines?".This tidy data set contains 1,599 red 
wines with 11 variables on the chemical properties of the wine. At least 3 
wine experts rated the quality of each wine, providing a rating between 0 
(very bad) and 10 (very excellent). This data set is a tidy set hence there 
is no need to clean it. It contains some important characteristics of red 
wines such as its pH, density, alcohol% present etc.

# Univariate Plots Section

```{r echo=FALSE,message=FALSE, warning=FALSE, Univariate_Plots}
#qplot(x = density, data = wr)
qplot(x = pH, data = wr)
#qplot(x = alcohol, data = wr)
#qplot(x = quality, data = wr)
#qplot(x = citric.acid, data = wr)
#qplot(x = residual.sugar, data = wr)
summary(wr$pH)
#Reason: To study univariate analysis of each variable in the dataset.
```

Reason: To study univariate analysis of each variable in the dataset.
Reflection: 'pH' is plotted on X-axis. And from below all the rest variables 
have been plotted on X-axis for studying them.

```{r echo=FALSE,message=FALSE, warning=FALSE}
qplot(x = density, data = wr)
```

```{r echo=FALSE,message=FALSE, warning=FALSE}
qplot(x = alcohol, data = wr)
```

The below plot is shown by eliminating the outliers.

```{r echo=FALSE,message=FALSE, warning=FALSE}
qplot(x = alcohol, data = wr) + xlim(8.5,14)
```

```{r echo=FALSE,message=FALSE, warning=FALSE}
#qplot(x = quality, data = wr)
ggplot(wr, aes(x=factor(quality))) + geom_bar()
```

```{r echo=FALSE,message=FALSE, warning=FALSE}
qplot(x = citric.acid, data = wr)
```

The below plot is shown by eliminating the outliers.

```{r echo=FALSE,message=FALSE, warning=FALSE}
qplot(x = citric.acid, data = wr) + xlim(0.00,0.75)
```

```{r echo=FALSE,message=FALSE, warning=FALSE}
qplot(x = residual.sugar, data = wr)
```

The below plot is shown by eliminating the outliers.

```{r echo=FALSE,message=FALSE, warning=FALSE}
qplot(x = residual.sugar, data = wr) + xlim(1,5)
```

```{r echo=FALSE,message=FALSE, warning=FALSE}
qplot(x = fixed.acidity, data = wr)
```

The below plot is shown by eliminating the outliers.

```{r echo=FALSE,message=FALSE, warning=FALSE}
qplot(x = fixed.acidity, data = wr) + xlim(5,13)
```

```{r echo=FALSE,message=FALSE, warning=FALSE}
qplot(x = volatile.acidity, data = wr)
```

The below plot is shown by eliminating the outliers.

```{r echo=FALSE,message=FALSE, warning=FALSE}
qplot(x = volatile.acidity, data = wr) + xlim(0.2,1.1)
```

```{r echo=FALSE,message=FALSE, warning=FALSE}
qplot(x = sulphates, data = wr) 
```

The below plot is shown by eliminating the outliers.

```{r echo=FALSE,message=FALSE, warning=FALSE}
qplot(x = sulphates, data = wr) + xlim(0.4,1.1)
```

```{r echo=FALSE,message=FALSE, warning=FALSE}
qplot(x = free.sulfur.dioxide, data = wr) 
```

The below plot is shown by eliminating the outliers.

```{r echo=FALSE,message=FALSE, warning=FALSE}
qplot(x = free.sulfur.dioxide, data = wr) + xlim(0,45)
```

```{r echo=FALSE,message=FALSE, warning=FALSE}
qplot(x = total.sulfur.dioxide, data = wr) 
```

The below plot is shown by eliminating the outliers.

```{r echo=FALSE,message=FALSE, warning=FALSE}
qplot(x = total.sulfur.dioxide, data = wr) + xlim(0,150)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
qplot(x = pH, data = wr) 
```

The below plot is shown by eliminating the outliers.

```{r echo=FALSE, message=FALSE, warning=FALSE}
qplot(x = pH, data = wr, xlim = c(3.0, 3.75)) 
#Reason: To get better review of the histogram.
```

Reason: To get better review of the histogram.
Reflection: The X-axis has been limited to pH 3.0 to 3.75 for better 
understanding of the Histogram.

```{r echo=FALSE, message=FALSE, warning=FALSE}

p1 <- qplot(x = alcohol, data = wr)
p2 <- qplot(x = log10(alcohol + 1), data = wr)
p3 <- qplot(x = sqrt(alcohol), data = wr)

grid.arrange(p1, p2, p3, ncol = 1)
#Reason: To find if there are any variations by using logarithmic and sqrt 
#functions
```


Reason: To find if there are any variations by using logarithmic and sqrt 
functions
Reflection: Here, we cannot see much difference in the histograms hence there is
no need to transform data, we also need not compare any two features, so the
box-plots haven't been used!

# Univariate Analysis

### What is the structure of your dataset?

  The Structure of this dataset is a dataframe

### What is/are the main feature(s) of interest in your dataset?

  The main features of interest in this dataset are, 'density', 'pH',
'alcohol', 'quality'

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

  'fixed.acidity', 'residual.sugar', 'citric.acid' such other features
will help support the investigation into the features of interest. 

### Did you create any new variables from existing variables in the dataset?

  There was no need here to create any new variables from existing 
variables in the dataset. 

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the 
form \ of the data? If so, why did you do this?

  No, there were not any unsual distributions seen in the features investigated.
Also, there was no need to perform any operations for data munging or tidying 
of data as this dataset is already to structured, tidy and easy to work with.

# Bivariate Plots Section

 Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}
qplot(pH, density, data = wr)
```

Reason: The below plot can be properly studied as compared to the above one.
Reflection:   density is plotted against pH in the above and below plots.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = pH, y = density), data = wr) + geom_point(alpha = 0.3) + 
  xlim(3.0,3.6) + ylim(0.992, 1.002)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = fixed.acidity, y = volatile.acidity), data = wr) + 
  geom_point(alpha = 1/2, position = position_jitter(h = 0)) + 
  xlim(4, 13) + coord_trans(y = "sqrt") 
```

Reason: We can check if there is any relation between these two.
Reflection:   fixed acdity and volatile acidity with square root of volatile 
acidity are plotted

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide), data = wr) + 
  geom_point(alpha = 1/5) + xlim(0,40) + ylim(0,150)

```

Reason: To get to know the relation between free sulfur dioxide and total sulfur
dioxide. We come to know how much one is dependent on other.
Reflection: scatterplot of free sulfur dioxide and total sulfur dioxide is seen

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = sulphates, y = total.sulfur.dioxide), data = wr) + 
  geom_point(alpha = 1/3, color = "orange") + xlim(0.4, 0.9) + ylim(0,125) +
  geom_line(stat = 'summary',fun.y = mean, color = "red") 
```

The mean of total.sulfur.dioxide against sulphates is seen in this 
plot. This plot the variation between the total.sulfur.dioxide and sulphates 
hence I have chosen this plot. Also I found it as interesting variation even
though these two are not the main features of interest.

```{r echo=FALSE, message=FALSE, warning=FALSE}
wr.qual_by_alcohol <- wr %>%
  group_by(alcohol) %>%
  summarise(mean_quality = mean(quality),
          median_quality = median(quality),
          n = n()) %>%
  arrange(alcohol)
head(wr.qual_by_alcohol)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(alcohol, mean_quality), data = wr.qual_by_alcohol) + 
  geom_line()
#Reason: To see how much the quality varies alongwith the alcohol
```

Reason: To see how much the quality varies alongwith the alcohol
Reflection: We see a downfall with wines containing 10% alcohol in them.
hmm...interesting!


```{r echo=FALSE, message=FALSE, warning=FALSE}
with(wr, cor.test(pH, quality, method = "pearson"))

with(wr, cor.test(alcohol, quality, method = "pearson"))

with(wr, cor.test(density, quality, method = "pearson"))

#with(wr, cor.test(density, alcohol, method = "pearson"))
```
Reason: to find the strongest relationship between two characteristics.
Reflection: Correaltion test is performed to check how strong the relationship 
is between two characteristics. Alcohol and quality is the highest.


# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features 
in \ the dataset?

 Relation between Alcohol and quality is the strongest realtionship 
seen.We can see that majorly the quality of Red wines depends upon the 
alcohol percentage in them. Then comes the acidity and pH and then density.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

 (1)free.sulfur.dioxide and total.sulfur.dioxide (2)fixed.acidity 
and volatile.acidity. These two relationships would catch interest which are
other than main features we need. 

### What was the strongest relationship you found?

 The realtionship between Alcohol and quality is the strongest relationship 
 seen in the main features of interest because the Pearson's coeffiecient is 
 the highest (0.476)

# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}
wr.qual_by_alcohol_density <- wr %>%
  group_by(alcohol, density) %>%
  summarise(mean_quality = mean(quality),
            median_quality = median(quality),
            n = n()) %>%
  ungroup() %>%
  arrange(alcohol)

head(wr.qual_by_alcohol_density)
#Reason: To study multivariate plots a new dataframe is introduced.
```

Reason: To study multivariate plots a new dataframe is introduced.
Reflection:   A dataframe is made containg the quality with alcohol and density 
for easy comparision.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = density, y = alcohol),
       data = wr.qual_by_alcohol_density) + geom_point(alpha = 1/2, 
                                                       color = 'red')
```



# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
 investigation. Were there features that strengthened each other in terms of
 or looking at your feature(s) of interest?

 There was not much strengthening of relationships as compared to the
bivariate analysis. Wines containing around 8%-12% of alcohol and 
0.995-1.000g/cc denser are the better types of wines!

### Were there any interesting or surprising interactions between features?

 To my suprise, the multivariate plots did not show much interesting relations 
 compared to the bivariate plots. 

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
ggplot(aes(x = pH, y = density), data = wr) + 
  geom_point(alpha = 0.2, color = 'blue') + 
  xlim(3.0,3.75) + 
  ggtitle("Density(g/cm^3) against pH")
```

### Description One
 This plot shows that the wines which are acidic with pH around 
3.2-3.4 and denser 0.995g/cc-1.000g/cc are better wines. This plot gives us 
a good idea of the characteristics of Wines hence I have chosen this plot. 

### Plot Two

```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}


ggplot(aes(x = sulphates, y = total.sulfur.dioxide), data = wr) + 
  geom_point(alpha = 1/3,position = position_jitter(h = 0),
             color = "orange") + 
  xlim(0.4, 0.9) + ylim(0,125) +
  geom_line(stat = 'summary',fun.y = mean, color = "red") +
  ggtitle("Mean total Sulfur dioxide in Wines.total.sulfur.dioxide(mg/dm^3)
          vs Sulphates(g/dm^3)")


```


### Description Two

The red line which we see is the mean line for total.sulfur.dioxide. The mean of 
total.sulfur.dioxide against sulphates is seen in this 
plot. This plot the variation between the total.sulfur.dioxide and sulphates 
hence I have chosen this plot. Also I found it as interesting variation even
though these two are not the main features of interest.


### Plot Three

```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}
ggplot(aes(x = density, y = alcohol),
       data = wr.qual_by_alcohol_density) + 
  geom_point(alpha = 1/2, color = 'red') +
  ggtitle("Red Wines-Alcohol(%) vs. Density(g/cm^3)")


```

### Description Three

 The above scatterplot shows that the red wines with density 
0.995g/cc to 1.000g/cc containing 8%-12% alcohol are better. This plot gives a
better/clear idea of the characteristics hence I have chosen this plot. 

------

# Reflection

 The dataset I chose to explore is the Red Wine Quality. I chose this dataset 
just out of interest :) ! This dataset was already structured as a 
result I did not need to restructure it or data munging was not necessary. 
I found that the wines containing alcohol 8%-12%, density 0.995g/cc-1.000g/cc 
and pH 3.2-3.4 are better wines.While exploring, my first hurdle was to choose
the correct characteristics to work on and explore. Then when I started 
plotting, as I went on exploring I found it easy to work with. Plotting the
bivariate plotting was helpful more than the multivariate plotting. I also 
found an interesting realtionship between total.sulfur.dioxide and sulphates 
present in the wines. We can add more data in the present dataset of the 
present wines and hence work on the same modified dataset to check in the 
future if there are any changes in the current findings and the reason 
behind the changes in this time period!
